Tracing the cultural evolution of meaning at the semantics-pragmatics interface

### 2017-03-20 paper/paper-v3.tex
rev
plots rev
revision draft

We begin setting the stage by defining the learning prior and the set of types under consideration. Drawing from our preceding discussion, functional pressure on successful communication combined with learning pressures in the form of a bias against upper-bounds may lead to the selection of $\Llack$-like semantics and the majority view of a division of labor between semantics and pragmatics. Before coming to this, it is instructive to first inspect the effect of these pressures in isolation. We therefore first showcase the effects that expressivity and learnability have on their own before turning to those resulting from their combination.

\subsubsection{An inductive learning bias in favor of semantic simplicity}
There is a growing effort to develop empirically  testable representational  languages that allow for the measure of semantic complexity. For instance, so-called {\em languages of thought} (LOT) have been put to test in various rational probabilistic models that show encouraging results (see, e.g., \citealt{katz+etal:2008, piantadosi+etal:underreview, piantadosi+etal:2012a} and \citealt{piantadosi+jacobs:2016} for recent discussion). At its core, a LOT defines a set of operations and composition rules from which lexical meaning can be derived. As a first approximation and for the sake of concreteness, we follow this approach to make the preference of learners for simpler semantic representations precise and define the complexity of a representation as a function of its derivation cost in a weighed generative LOT.

@@

where $Compl(c)$ is the complexity of the minimal derivation length of each concept. There are other more sophisticated ways to define priors over lexica (see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a}) but the key assumption here, common to all of them, is that simple representational expressions should be favored and complex ones disfavored. More generally, we should stress that these details -- from the generative grammar to its complexity measure -- are stipulated and are to be regarded as an illustratory instantiation of our general assumptions instead of as strong empirical claims.

@@

Out of these possible lexica, three kinds are of particular relevance. First, lexica that assign the same concept to more than one message. Such lexica lack in expressivity but are favored by the learning bias if they lexicalize simple concepts. Second, lexica that conventionalize upper-bounds to realize a (quasi-)partition of the state space. As discussed before in relation to $\Lbound$, users of these lexica need not resort to pragmatic reasoning to convey an upper-bound with weak scalar expressions. Instead, this information is already provided by the semantics of their lexicon, giving them an advantage in terms of expressivity. Types with such lexica are expected to be the main contenders to the natural language like semantics that do not lexicalize an upper-bound for weak scalar expressions. Third and lastly, $\Llack$-style lexica that, paired with mutual reasoning, can convey an upper-bound pragmatically. There are six lexica of the second kind and six of the third. The following three lexica exemplify each kind, with $\Lall$ being a lexicon of the first. This lexicon is not optimal for communication as it assigns all its messages to the same state, $\sall$. This inevitably leads to a communicative disadvantage in its use but it may nevertheless have a learnability advantage over our target $\Llack$ and its competitor $\Lbound$ in virtue of being the simplest lexicon complexity-wise, making it a priori favored by learners.

@@

Recall that types are a combination of a lexicon and a manner of language use. We analyze the model's predictions in populations of types with one of the two signaling behaviors introduced earlier; literal or pragmatic. The former correspond to level $0$ reasoners and the latter to ones of level $1$. Higher level reasoning is not required to derive scalar implicatures from the lexica we consider here, nor do they leave room for substantial pragmatic refinement. Accordingly, there are $432$ types in this model. Six are variants of our pragmatic $\Llack$ target and twelve are either literal or pragmatic speakers of variants of its main contender $\Lbound$.

@@

\subsubsection{Expressivity only}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/fig1-onlyr}
\caption{Stacked proportion of pragmatic $\Llack$-style types and incumbent types, when other than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under only a pressure for expressivity. The right-most plots zoom in on only the proportion of pragmatic $\Llack$ types.}
\label{fig:only-R}
\end{figure}

Communicative efficiency is sensitive to $\lambda$ as it influences signaling behavior. It therefore has a central bearing on types' fitness. This is showcased in Figure \ref{fig:only-R}, which depicts the proportion of pragmatic $\Llack$-style speakers in $5$ independent populations after $50$ generations, as well as that of the type with the highest proportion in a given population, the incumbent. As is to be expected, less expressive types do not fare well under pressure for communicative success and those able to convey states less equivocally fare better. This difference increases with $\lambda$ as it promotes more rational linguistic behavior, widening the gap in expressivity between some types. As suggested by Figure \ref{fig:only-R} the incumbent in most populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure only for expressivity does not lend a justification to the systematic prevalence of a division of labor between semantics and pragmatics that we set out to explain, under any $\lambda$ value. For instance, in $1000$ independent populations with $\lambda = 20$ pragmatic $\Llack$ types only make up $11$ cases of incumbency, corresponding to a mean proportion of $0.003$ across populations. By contrast, $913$ incumbents were $\Lbound$-style users with close to an even share between literal ($454$) and pragmatic types ($459$), corresponding to a mean proportion of about $.48$ taken together. Across types and populations incumbents make up $.50$ of the mean outcome.

@@

The effect of iterated learning without a pressure for expressivity with posterior sampling ($l = 1$) and a stronger tendency towards posterior maximization ($l = 15$) is shown in Figure \ref{fig:only-M} together with the prior over types. The prior shows that while users of $\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are nevertheless more advantaged than others such as $\Lbound$ in virtue of the relatively simple semantics they conventionalize. Crucially, $\Llack$ enables its users to convey each state with a single message when combined with pragmatic reasoning provided sufficiently high $\lambda$. This makes it less likely to be confused with other types if the learning data is not too sparse ($k \geq 5$). In sum, learners have a higher propensity to infer pragmatic $\Llack$ over types that might be hard or impossible to distinguish it from based on overt linguistic behavior, such as $\Lbound$, and $\Llack$ is also less likely to be mistaken with types with different linguistic behavior because its pragmatic use approximates a one-to-one form-meaning mapping. As a consequence, a stronger propensity to maximize the posterior increases their proportion in the population.

In contrast to a pressure only for expressivity with higher $\lambda$, learnability does not succeed in selecting for a single type of a kind. For example, only one of the six pragmatic $\Lllack$ types. Instead, it leads to their coexistence. Each is passed on to next generations with the same faithfulness and, differently from a pressure for communicative success within a population, they do not stand in competition with each other. Nevertheless, as implicitly suggested by Figure \ref{fig:only-M}, in $1000$ independent populations all incumbents were pragmatic $\Llack$ users provided sufficient rationality in linguistic behavior. Low values of $\lambda$ instead level the differences in linguistic behavior across types and therefore reflect the learners' inductive bias more faithfully, favoring functionally deficient but a priori preferred types such as literal $\Lall$. A pressure for learnability alone may consequently lead to the selection of communicatively suboptimal types that are easier to learn. In the extreme, when $l = 1$ and $\lambda = 1$ all of $1000$ independent populations had $\Lall$ as incumbent.

\subsubsection{Expressivity and learnability.}
Inspecting these pressures separately not only showcases the contribution of the model's components but also highlights some of their broader implications. First and foremost, neither dynamic on its own comes close to converging to a monomorphic population of pragmatic $\Llack$ speakers. On the one hand, the slight communicative advantage of $\Lbound$'s lexicalization of an upper-bound over pragmatic use of $\Llack$ leads to the former's selection under a pressure for expressivity. On the other hand, pressure only towards learnability has a modest but clear effect of promoting pragmatic $\Llack$ over functionally similar but semantically more complex alternatives such as $\Lbound$. However, on its own learnability does not foment the propagation of these types across the population. Furthermore, this pressure alone leads to heterogeneous populations that do not justify the selection of a particular type over others.

@@

Figure \ref{fig:rmd} illustrates the combined effects of both pressures for a sample of $\lambda$ and $l$ values. These results show that an inductive learning bias for simpler semantics in tandem with functional pressure can lead to the selection of $\Llack$-like semantics over $\Lbound$. As when each pressure was considered in isolation, the proportion of these types increases with $\lambda$ and $l$. Crucially, the addition of a pressure towards expressivity magnifies the learning outcome and dampens the proliferation of similar or identical types. As stressed above, this suggests that neither a learning bias nor functional pressure alone but their combination may lead to the lack of upper-bounds in the lexical meaning of scalar expressions in a mostly monomorphic population.

As before, low $\lambda$ and $l$ lead to the incumbency of communicatively suboptimal types that are a priori favored by learning, such as $\Lall$. An increase in $\lambda$ leads to the selection and incumbency of pragmatic $\Llack$ types but does not lead to monomorphic populations if learners sample from the posterior. Finally, a combination of sufficiently rational signaling behavior with some tendency for posterior maximization leads to increasingly monomorphic populations of types that convey upper-bounds pragmatically. A more succinct representation of the difference between the mean of the highest pragmatic $\Llack$-like types and the highest proportion of other types across $\lambda$ and $l$ values is given in Figure \ref{fig:diff}. \tb{Say more}

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=6cm,keepaspectratio]{./plots/fig4-incumbents-difference}
\caption{Difference between mean proportion of highest pragmatic $\Llack$ type and highest other type in $1000$ independent populations after $50$ generations under both pressures ($k = 5$).}
\label{fig:diff}
\end{figure}

@@

To recapitulate, other than the involvement of both pressures, the resulting proportion of pragmatic $\Llack$ speakers primarily hinges on three factors. First, the degree to which linguistic behavior is deterministic. This plays a role both in expressivity as well as in producing data that allows learners to discriminate this type from others. Second, the inductive bias, which leads learners to prefer simpler over more complex semantic representations in acquisition. Lastly, the posterior parameter, as it magnifies the effects of the learning bias in tandem with replication.

@@

A lack of upper-bounded in the lexical meaning of weak scalar expressions constitutes the majority view in the literature. However, it not clear to what extent other types should be present in the final population. It seems reasonable to expect functionally suboptimal lexica such as $\Lall$ to be ruled out because they fail to enable their users to communicate effectively and are not evinced in natural language. However, this is not true of $\Lbound$. The prediction that natural language communities are homogeneous or that a single speaker may entertain $\Llack$-like semantics for one scalar expression and $\Lbound$-like semantics for another is not implausible (cf. \citealt{franke+degen:2016}). Alternatively, a stronger tendency for posterior maximization in learning has to be assumed. This empirical issue relates to other two aspects left undiscussed: disadvantages of pragmatic reasoning and the effect of state frequencies on the fossilization of pragmatic inferences. We tacitly assumed pragmatic reasoning to come at no cost. However, there is experimental evidence that suggests  that the pragmatic derivation of upper-bounds costs effort and takes additional processing time (cf. \citealt{deNeys+schaeken:2007, huang+snedeker:2009}). This raises the question at which point such usage-based cost undercuts the learnability advantage of simpler semantic representations. Should cost play a role, then its effect is bound to depend on the frequency with which a given scalar expression is used. It may be that frequently drawn scalar implicatures fossilize to avoid cost whereas infrequent ones are still derived on-line. This also opens a possible venue to address the preceding question about the expected presence of $\Lbound$-like semantics in a population, but further empirical evidence is needed to assess these matters.

\section{General discussion}\label{sec:discussion}
We laid out a model that combines game theoretical models of functional pressure towards efficient communication \citep{nowak+krakauer:1999}, effects of transmission perturbations on (iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012, franke+jaeger:2014} as well as different lexica \citep{bergen+etal:2012,bergen+etal:2016}, and \tb{mention LOT here}. This model generates predictions about lexicalization patterns and, more generally, effects of communicative pressures on the cultural evolution of language.

We argued that the puzzle raised by semantics in light of pragmatics is hard to explain on purely functional grounds and that part of the answer may instead lie in the way transmission shapes the outcome of cultural evolution in tandem with a pressure for successful information transfer. In the realm of inductive biases, we adopted the assumption that simpler semantic representations are more likely to be learned (cf. \citealt{chater+vitanyi:2003}), \tb{mention LOT here}. Under this view, semantics and pragmatics play a synergic role in that representational simplicity is supplemented by pragmatic reasoning to counteract functional disadvantages otherwise incurred. As a consequence, iterated transmission and use of language lead to a regularization that may explain the lack of lexicalization of systematic pragmatic enrichments. This result is of particular relevance for the longstanding assumption of a divide and interaction between semantics and pragmatics by offering an account of why certain pragmatic inferences fail to lexicalize.

The main innovations of the model are its modular separation of expressivity and learnability, allowing for their isolated and combined analysis, the learning process involving a joint inference over types of pragmatic behavior and lexical meaning. The goal to decouple but model both expressivity and learnability has also recently been addressed by \citet{kirby+etal:2015}. In contrast to our proposal, Kirby et al. model expressivity as exerting its force only in the production of learning data. On a general level, this model's expressivity parameter has a similar effect to $\lambda$ in that it can give a communicative advantage to unambiguous languages over ambiguous ones. \tb{Should we go into more detail here? I think it's fine if we leave it out but $\lambda$ has a broad motivation whereas Kirby et al.'s parameter is quite explicitly a penalty parameter against ambiguity. I mainly wanted to give an idea of what the pressure for expressivity in their model does rather than argue for why ours is more principled.} However, the degree of mutual understanding of interlocutors central to replication and to our notion of expressivity is not taken into consideration. That is, while our proposal combines bidirectional horizontal transmission with its vertical and unidirectional counterpart, Kirby et al.'s model only considers the latter's influence. Our reasoning behind the inclusion of the former lies in the empirical and theoretical observation that learnability alone can lead the selection of functionally defective languages, as discussed in relation to $\Lall$ when $\lambda$ or $k$ are low. This outcome has been reported in a number of laboratory experiments where subjects learned and subsequently reproduced the language produced by a previous participant over multiple iterations, leading to a proliferation of ambiguous languages that associate a large number of meanings with a single form (see, e.g., \citealt{silvey+etal:2014} and experiment 1 in \citealt{kirby+etal:2008}). In contrast, experiments involving an interactive component have been found to foster languages that enable interlocutors to communicatively distinguish meanings more accurately  (e.g., \citealt{fay+etal:2013}; for a review of laboratory results under the iterated learning paradigm and further discussion see \citealt{kirby+etal:2015, tamariz+kirby:2016}). It is not evident how to compare these empirical findings given that they consider distinct meaning spaces, modes of transmission, iterations and feedback given to participants. However, we take these results to suggest that there is an important difference between a language generating learnable linguistic data and its actual performance as a means of information transfer. \tb{Explain the next two sentences} The former solely depends on the mechanism by which speakers associate form and meaning. The latter additionally hinges on the addressee's linguistic experience and her ability to interpret linguistic input based on this experience. In sum, we contend that successful information transfer in a linguistic community is central to the adoption of a communication system and that this measure is not adequately reflected by production alone.

\section{Conclusion}
The cultural evolution of language is influenced by intertwined pressures. We set out to investigate this process by putting forward a model that combines a pressure toward efficient and successful information transfer with perturbations that may arise in the transmission of linguistic knowledge in acquisition. \tb{Clarify and put this in better words. This is quite central:} Additionally, we argued for the necessity of considering the role of pragmatics in investigations on the cultural evolution of meaning. These components and their mutual influence were highlighted in a case study on the lack of lexical upper-bounds in weak scalar expressions that showed that, when pressured for learnability and expressivity, the former drives for simpler semantic representations inasmuch as pragmatics can compensate for lack of expressivity in use. That is, the relative learning advantage of simpler semantics in combination with  functional pressure in use may offer an answer to why natural languages fail to lexicalize systematic pragmatic inferences.

>>>>

We begin setting the stage by defining the learning prior and the set of types under consideration. Drawing from our preceding discussion, functional pressure on successful communication combined with learning pressures in the form of a bias that favors simpler semantics may lead to the selection of $\Llack$-like semantics and to the majority view of a division of labor between semantics and pragmatics. We first showcase the effects that expressivity and learnability have in isolation before turning to those resulting from their combination.

\subsubsection{An inductive learning bias for semantic simplicity}
There is a growing effort to develop empirically  testable representational  languages that allow for the measure of semantic complexity. For instance, so-called {\em languages of thought} (LOT) have been put to test in various rational probabilistic models that show encouraging results (see, e.g., \citealt{katz+etal:2008, piantadosi+etal:underreview, piantadosi+etal:2012a} and \citealt{piantadosi+jacobs:2016} for recent discussion). At its core, a LOT defines a set of operations and composition rules from which lexical meaning can be derived. As a first approximation and for the sake of concreteness, we follow this approach to make the preference of learners for simpler semantic representations precise and define the complexity of a representation as a function of its derivation cost in a weighed generative LOT.

@@

where $Compl(c)$ is the complexity of the minimal derivation length of $c$ according to the grammar. There are many ways to define priors over lexica (see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a}) but the key assumption here, common to all of them, is that simple representational expressions should be favored over more complex ones. We should stress that these details -- from the generative grammar to its complexity measure -- are to be regarded as an instantiation of our general assumptions instead of as empirical claims in their own right.

@@

Out of these possible lexica, three kinds are of particular relevance. First, lexica that assign the same concept to more than one message. Such lexica lack in expressivity but are favored by the learning bias should they lexicalize simple concepts. Second, lexica that conventionalize upper-bounds to realize a (quasi-)partition of the state space. As discussed before in relation to $\Lbound$, users of these lexica need not resort to pragmatic reasoning to convey an upper-bound with weak scalar expressions. Instead, this information is already provided by the semantics of their lexicon, giving them an advantage in terms of expressivity. Types with such lexica are expected to be the main contenders of the natural language like semantics that do not lexicalize an upper-bound for weak scalar expressions. Third and lastly, $\Llack$-style lexica that, paired with mutual reasoning, can convey an upper-bound pragmatically. There are six lexica of the second kind and six of the third. The following three lexica exemplify each kind, with $\Lall$ being a lexicon of the first kind. This lexicon is not optimal for communication as it assigns all its messages to the same state, $\sall$. This inevitably leads to a communicative disadvantage in its use. However, it may have a learnability advantage over our target $\Llack$ and its competitor $\Lbound$ in virtue of being the simplest lexicon, making it a priori favored by learners.

@@

Recall that types are a combination of a lexicon and a manner of language use. We analyze the model's predictions in populations of types with one of the two behaviors introduced earlier; literal or pragmatic. The former correspond to level $0$ reasoners and the latter to ones of level $1$. Higher level reasoning is not required to derive scalar implicatures from these lexica, nor do they leave room for substantial pragmatic refinement. Accordingly, we consider a total of $432$ types. Six are variants of our pragmatic $\Llack$ target and twelve are either literal or pragmatic speakers of variants of its main contender, $\Lbound$.

@@

\subsubsection{Expressivity only} 
\begin{figure}
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/fig1-onlyr}
\caption{Stacked proportion of pragmatic $\Llack$-style types and incumbent types, when other than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under only a pressure for expressivity. The right-most plots zoom in on only the proportion of pragmatic $\Llack$ types.}
\label{fig:only-R}
\end{figure}

Communicative efficiency is sensitive to $\lambda$ as it influences signaling behavior. This is showcased in Figure \ref{fig:only-R}, which depicts the proportion of pragmatic $\Llack$-style speakers in $5$ independent populations after $50$ generations, as well as that of the type with the highest proportion in a given population, the incumbent. As is to be expected, less expressive types do not fare well under pressure for communicative success and those able to convey states less equivocally fare better. This difference increases with $\lambda$ as it promotes more rational linguistic behavior, widening the gap in expressivity between some types. As suggested by Figure \ref{fig:only-R}, the incumbent in most populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure only for expressivity does not lend a justification to the systematic prevalence of a division of labor between semantics and pragmatics under any $\lambda$-value. For instance, in $1000$ independent populations with $\lambda = 20$ pragmatic $\Llack$ types only make up $11$ cases of incumbency, corresponding to a mean proportion of $0.003$ across populations. By contrast, $913$ incumbents were $\Lbound$-style users with close to an even share between literal ($454$) and pragmatic types ($459$), corresponding to a mean proportion of about $.48$ taken together. %Across types and populations incumbents make up $.50$ of the mean outcome.

@@

The effect of iterated learning without a pressure for expressivity using either posterior sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$) is shown in Figure \ref{fig:only-M} together with the prior over types. The prior shows that while users of $\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are nevertheless more advantaged than others such as $\Lbound$ in virtue of the relatively simple semantics they conventionalize. Crucially, $\Llack$ enables its users to convey each state with a single message when combined with pragmatic reasoning provided sufficiently high $\lambda$. This makes it less likely to be confused with other types if the learning data is not too sparse ($k \geq 5$). Put differently, learners have a higher propensity to infer pragmatic $\Llack$ over types that might be hard or impossible to distinguish it from based on overt linguistic behavior, such as $\Lbound$, and $\Llack$ is also less likely to be mistaken with types with different observable behavior because its pragmatic use approximates a one-to-one form-meaning mapping. As a consequence, a stronger propensity to maximize the posterior increases their proportion in the population.

However, in contrast to a pressure only for expressivity with high $\lambda$, learnability does not succeed in selecting for a single type of a kind. For instance, for only one of the six pragmatic $\Llack$ types. Instead, it leads to their coexistence. Each is passed on to next generations with the same faithfulness and, differently from a pressure for communicative success, they do not stand in competition with each other.

As implicitly suggested by Figure \ref{fig:only-M}, in $1000$ independent populations all incumbents were pragmatic $\Llack$ users provided sufficient rationality in linguistic behavior, with each reaching approximately the same proportion of users in the population. Low values of $\lambda$ level the differences in observable behavior across types and therefore reflect the learners' inductive bias more faithfully, favoring functionally deficient but a priori preferred types such as literal $\Lall$. A pressure for learnability alone may consequently lead to the selection of communicatively suboptimal types that are easier to learn. In the extreme, when $l = 1$ and $\lambda = 1$ all of $1000$ independent populations had users of $\Lall$ as incumbents.

\subsubsection{Expressivity and learnability}
The separate inspection of each pressure not only showcases the contribution of the model's components but also highlights some of their broader implications for a division of labor between semantics and pragmatics. Neither dynamic on its own comes close to converging to a monomorphic population of pragmatic $\Llack$ speakers. When pressured toward expressivity, the slight communicative advantage of $\Lbound$'s lexicalization of an upper-bound over pragmatic use of $\Llack$ leads to the former's selection. When pressured toward learnability, pragmatic $\Llack$ is promoted over functionally similar but semantically more complex alternatives such as $\Lbound$. However, on its own learnability does not foment the propagation of these types across the population. What is more, it leads to heterogeneous populations that foster an equal share of types of a kind, such as different pragmatic $\Llack$ types.

@@

Figure \ref{fig:rmd} illustrates the combined effects of both pressures for a sample of $\lambda$ and $l$ values. These results show that an inductive learning bias for simpler semantics in tandem with functional pressure can lead to the selection of $\Llack$-like semantics paired with pragmatic language use. The proportion of these types increases with $\lambda$ and $l$. Pressure towards expressivity magnifies the effects of iterated learning and dampens the proliferation of types of a kind that are equal in expressivity {\em and} learnability. A pressure towards learnability favors the transmission of simpler semantics and thereby makes types that exploit pragmatic reasoning to strengthen meaning more likely to be adopted over alternatives that do not. In short, neither learnability nor functional pressure alone but their combination may lead to the lack of upper-bounds in the lexical meaning of scalar expressions in predominantly monomorphic populations.

As before, low $\lambda$ and $l$ lead to the incumbency of communicatively suboptimal types that are a priori favored, such as $\Lall$. An increase in $\lambda$ leads to the selection and incumbency of pragmatic $\Llack$ but does not lead to monomorphic populations if learners sample from the posterior. Finally, a combination of sufficiently rational signaling behavior with some tendency for posterior maximization leads to increasingly monomorphic populations of types that convey upper-bounds pragmatically. The difference between the mean of the highest pragmatic $\Llack$-like type in $1000$ independent populations and the highest proportion of other types across $\lambda$ and $l$ values is shown in Figure \ref{fig:diff}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=6cm,keepaspectratio]{./plots/fig4-incumbents-difference}
\caption{Difference between mean proportion of highest pragmatic $\Llack$ type and highest other type in $1000$ independent populations after $50$ generations under both pressures ($k = 5$). \tb{I will run more simulations with higher $\lambda$ and $l$. I'm pretty sure the difference still increases.}}
\label{fig:diff}
\end{figure}

@@

To recapitulate, other than the involvement of both pressures, the resulting proportion of pragmatic $\Llack$ speakers primarily hinges on three factors. First, the degree to which linguistic behavior is rational in that agents try to maximize communicative success. This plays a role both in expressivity as well as in producing data that allows learners to discriminate this type from others. Second, the inductive bias, which leads learners to prefer simpler over more complex semantic representations in acquisition. Lastly, the posterior parameter, as it magnifies the effects of the learning bias in tandem with replication.

@@

A lack of upper-bounds in the lexical meaning of weak scalar expressions constitutes the majority view in the literature. However, it not clear to what extent other types should be present in the final population. It seems reasonable to expect functionally suboptimal lexica such as $\Lall$ to be ruled out but matters are different for $\Lbound$. The prediction that natural language communities are homogeneous or that speakers may entertain $\Llack$-like semantics for one scalar expression and $\Lbound$-like semantics for another is not implausible (cf. \citealt{franke+degen:2016}). Alternatively, a stronger tendency for posterior maximization in learning has to be assumed. This empirical issue relates to other two aspects left undiscussed: disadvantages of pragmatic reasoning and the effect of state frequencies on the fossilization of pragmatic inferences. We tacitly assumed pragmatic reasoning to come at no cost. However, there is experimental evidence that suggests  that the pragmatic derivation of upper-bounds costs effort and takes additional processing time (cf. \citealt{deNeys+schaeken:2007, huang+snedeker:2009}). This raises the question at which point such usage-based cost undercuts the learnability advantage of simpler semantic representations. Should cost play a role, then its effect is bound to depend on the frequency with which a given scalar expression is used. It may be that frequently drawn scalar implicatures fossilize to avoid cost whereas infrequent ones are derived on-line to avoid complexity in acquisition. This also opens a possible venue to address the preceding question about the expected presence of $\Lbound$-like semantics in a population, but further empirical evidence is needed to assess these matters.

\section{General discussion}\label{sec:discussion}
We laid out a model that combines game theoretical models of functional pressure towards efficient communication \citep{nowak+krakauer:1999}, effects of transmission perturbations on (iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012, franke+jaeger:2014} as well as different lexica \citep{bergen+etal:2012,bergen+etal:2016}. This model generates predictions about the co-evolution of conventional meaning and pragmatic reasoning types and, more generally, effects of communicative pressures on the cultural evolution of language.

We argued that the puzzle raised by semantics in light of pragmatics is hard to explain on purely functional grounds and that part of the answer may instead lie in the way successful transmission of linguistic knowledge across and within populations shapes the outcome of their co-evolution. In the realm of inductive biases, we adopted the assumption that simpler semantic representations are more likely to be learned (cf. \citealt{chater+vitanyi:2003}). Under this view, semantics and pragmatics play a synergic role in that representational simplicity is supplemented by pragmatic reasoning to counteract functional disadvantages otherwise incurred. As a consequence, iterated transmission and use of language lead to a regularization that may explain the lack of lexicalization of systematic pragmatic enrichments. This result is of particular relevance to the longstanding assumption of a divide and interaction between semantics and pragmatics by offering an account of why certain pragmatic inferences fail to lexicalize.

The main innovations of the model are its modular separation of expressivity and learnability, allowing for their isolated and combined analysis, as well as the learning process involving a joint inference over types of pragmatic behavior and lexical meaning. The goal to decouple but model both expressivity and learnability has also recently been addressed by \citet{kirby+etal:2015}. In contrast to our proposal, Kirby et al. model expressivity as exerting its force only in the production of learning data. The degree of mutual understanding of interlocutors central to replication and to our notion of expressivity is absent. That is, while our proposal combines bidirectional horizontal transmission with its vertical and unidirectional counterpart, Kirby et al.'s model only considers the latter's influence. Our reasoning behind the inclusion of the former lies in the empirical and theoretical observation that learnability alone can lead the selection of functionally defective languages, as showcased with $\Lall$. This outcome has been reported in a number of laboratory experiments where subjects learned and subsequently reproduced the language produced by a previous participant over multiple iterations, leading to a proliferation of ambiguous languages that associate many meanings with a single form (see, e.g., \citealt{silvey+etal:2014} and experiment 1 in \citealt{kirby+etal:2008}). In contrast, experiments involving an interactive component have been found to foster languages that enable interlocutors to communicatively distinguish meanings more accurately  (e.g., \citealt{fay+etal:2013}; for a review of laboratory results under the iterated learning paradigm and further discussion see \citealt{kirby+etal:2015, tamariz+kirby:2016}). It is not evident how to compare these empirical findings given that they consider distinct meaning spaces, modes of transmission, iterations and feedback given to participants. However, on a general level they may suggest that there is an important difference between a language generating learnable linguistic data and its actual performance as a means of information transfer. That is, we contend that successful information transfer in a linguistic community is central to the adoption of a communication system and that this measure is not adequately reflected by production alone.

\section{Conclusion}
The cultural evolution of meaning is influenced by intertwined pressures. We set out to investigate this process by putting forward a model that combines a pressure toward successful information transfer with perturbations that may arise in the transmission of linguistic knowledge in acquisition. Its objects of selection and replication are pairs of lexical meanings and patterns of language use. This allows the model to trace the interaction between conventional meaning and pragmatic use. Additionally, it takes the challenge serious of neither semantics nor pragmatics being directly observable. Instead, learners need to infer these unobservables from overt data that results from their combination.  These components and their mutual influence were highlighted in a case study on the lack of lexical upper-bounds in weak scalar expressions that showed that, when pressured for learnability and expressivity, the former force drives for simpler semantic representations inasmuch as pragmatics can compensate for lack of expressivity in use. That is, the relative learning advantage of simpler semantics in combination with functional pressure in use may offer an answer to why natural languages fail to lexicalize systematic pragmatic inferences.
